Expanding the scope of experimental archaeology using the Perception-Process-Product conceptual framework

### 2023-03-29 manuscript/manuscript.Rmd
added some references.

Experimental archaeology is based on the concept of analogy (i.e., the past is at least partially similar to the present in some aspects). It is acknowledged that the validity of this type of analogical inference has long been a subject of debate in archaeology (Wylie, 1985), and a comprehensive review of it is beyond the scope of this essay. For instance, presumably in the paleolithic period the development of stone knapping skills mostly happened during childhood and these children grew up in an environment surrounded by habitual stone tool users and makers. In this case, what can knapping teaching experiments involving modern adults who have zero exposure to stone tools inform us about the past learning behavior? It is important to clarify that no experimental project ever intends or claims to provide the perfect reconstruction of the past but rather aims at identifying variables relevant to the question of interest (Stout & Khreisheh, 2015), which can be often ignored through pure deductive reasoning. In the end, all experiments are wrong, but some are useful, and we need more of them.

@@

As implied in its name, the implementation of Triple P framework involves the collection of process-level (ethological) and perception-level (ethnographic) data (**Figure** \@ref(fig:concept)). Ethological approaches are widly used in the description and analysis of non-human animal behavior [@fragaszy2018]. new toolkit such as BORIS were introduced [@friard2016]

>>>>

Experimental archaeology is based on the concept of analogy (i.e., the past is at least partially similar to the present in some aspects). It is acknowledged that the validity of this type of analogical inference has long been a subject of debate in archaeology [@wylie1985; @chapman2016], and a comprehensive review of it is beyond the scope of this essay. For instance, presumably in the paleolithic period the development of stone knapping skills mostly happened during childhood and these children grew up in an environment surrounded by habitual stone tool users and makers. In this case, what can knapping teaching experiments involving modern adults who have zero exposure to stone tools inform us about the past learning behavior? It is important to clarify that no experimental project ever intends or claims to provide the perfect reconstruction of the past but rather aims at identifying variables relevant to the question of interest [@stoutSkillLearningHuman2015], which can be often ignored through pure deductive reasoning. In the end, all experiments are wrong, but some are useful, and we need more of them.

@@

As implied in its name, the implementation of Triple P framework involves the collection of process-level (ethological) and perception-level (ethnographic) data (**Figure** \@ref(fig:concept)). Ethological approaches are widely used in the description and analysis of non-human animal behavior [@fragaszy2018]. new toolkit such as BORIS were introduced [@friard2016]
